Exploring probability distributions for bivariate temporal granularities

### 2019-11-21 main.Rmd
rough structure for circular done - trash not removed

**Notation:  Hierarchy table** Let H: $(G, C, k, n)$ be a hierarchy model containing m linear granularities. $G_{i}$ represents the linear granularity of order i and $k_{i}$å represents the period length of $$G_{i}$$ with respect to $$G_{i+1}$$ and $C_{l,m}$ represents the circular granularity that relates linear granularity of order l and m, with $\forall l,m \in n, l<m$.

@@

\begin{equation} \label{eq1}
\begin{split}
C_{(x,y)}(z) & = C_{(x,x+1)}(z) + k(x, x+1)(C_{(x+1,y)}(z)-1) \\  
  & =   C_{(x,x+1)}(z) + k(x, x+1)[ C_{(x+1,x+2)}(z) + c(x+1, x+2)( C_{(x+2, y)}(z) - 1) - 1]\\
  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+1)c(x+1, x+2)(f(x+2, y) - 1)\\
  & =  f(x, x+1) + c(x, x+1)(f(x+1, x+2) - 1) + c(x, x+2)(f(x+2, y) - 1)\\
  &\vdots\\
  & = \sum_{i=0}^{order(y) - order(x) - 1} c(x, x+i)(f(x
  +i, x+i+1) - 1)\\
\end{split}
\end{equation}

<!-- some discussions on primary units of calendar -->
So far we have used the Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. Any civil day is divided into 24 hours and each hour into 60 minutes and each minute into 60 seconds. There are exceptions where in London, for example, length of each "hour" varies from about 39 minutes in December to 83 minutes in June. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:

@@

| Units	      | Conversion factor
|------------	|-------------------	|
|  kin      	| 20             	    |
|  uinal    	| 18            	    |
|  tun      	| 20           	      |
|  katun    	| 20            	    |
|  baktun   	| 1          	        |

Examples of multiple-order-up  granularities can be kin of the tun or kin of the baktun whereas examples of single-order-up granularities may include kin of the uinal, uinal of the tun etc.

>>>>

**Notation:  Hierarchy table** Let $H_n: (G, C, k)$ be a hierarchy model containing n linear granularities. $G_{l}$ represents the linear granularity of order l and $k_{l,m}$ represents the period length of $G_{l}$ with respect to $G_{m}$ and $C_{G(l),G(m)}$ represents the circular granularity that relates linear granularity of order l and m, $\forall l,m \in n, l<m$.

@@

\begin{equation} \label{eq3}
\begin{split}
C_{(G_l,G_m)}(z) & = C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1) \\  
  & =   C_{(G_l,G_(l+1)}(z) + k(l, l+1)[ C_{G_(l+1),G_(l+2)}(z) + k(l+1, l+2)( C_{G_(l+2), G_m)}(z) - 1) - 1]\\
  & =  C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{G_(l+1),G_(l+2)}(z) - 1) + k(l, l+1)k(l+1, l+2)(C(l+2, m) - 1)\\
  & =  C_{(G_l,G_(l+1)}(z) + k(l, l+1)(C_{G_(l+1),G_(l+2)}(z) - 1) + k(l, l+2)(C_{G_(l+2),G_(l+m)}(z) - 1)\\
  &\vdots\\
  & = \sum_{i=0}^{order(m) - order(l) - 1} k(l, l+i)(C_{G_(l+i),G_(l+i+1)}(z) - 1)\\
\end{split}
\end{equation}

<!-- some discussions on primary units of calendar -->
**Example:** So far we have used the Gregorian calendar as it is the most widely used calendar. But it is far from being the only one. All calendars fall under three types - solar, lunar or lunisolar/solilunar but the day is the basic unit of time underlying all calendars. Various calendars, however, use different conventions to structure days into larger units: weeks, months, years and cycle of years. The French revolutionary calendar divided each day into 10 "hours", each "hour" into 100 "minutes" and each "minute" into 100 "seconds". Nevertheless, for any calendar a hierarchy can be defined. For example, in Mayan calendar, one day was referred to as 1 kin and the calendar was structured as follows:

@@

|  G	        |C                 |k
|------------	|------------------|------------|
|  kin      	|kin-of-uinal      |  20        |
|  uinal    	|uinal-of-tun      |  18  	    |
|  tun      	|tun-of-katun      |  20	      |
|  katun    	|katun-of-baktun   |  20        |
|  baktun   	| 1                |	1         |

Examples of multiple-order-up  granularities can be kin-of-tun or kin-of-baktun whereas examples of single-order-up granularities may include kin-of-uinal, uinal-of-tun etc.

Let us use the equation \ref{eq3} to compute the multiple-order-up granularity uinal_katun for Mayan calendar, which is periodic.

\begin{equation} \label{eq4}
\begin{split}
C(uinal, baktun) & = C(uinal, tun) + kuinal, tun)C(tun,katun) + C(uinal, katun)C(katun, baktun) \\
              & = \lfloor z/20\rfloor \mod 18  + 20*\lfloor z/20*18\rfloor \mod 20 + 20*18*20\lfloor z/20*18*20\rfloor \mod 20 \\
\end{split}
\end{equation}

### Computation of single-order from multiple-order

For a hierarchy table $H_n: (G, C, k)$ with $l_1, l_2, m_1, m_2 \in {1, 2, \dots, n}$ and $l_2<l_1$ and $m_2>m_1$, we have

\begin{equation} \label{eq5}
C_{G_{l1}, G_{m1}}(z) = C_{G_{l2}, G_{m2}}(\lfloor z/k(l_2,l_1) \rfloor\mod k(m_1, m_2))
\end{equation}

**Example:** Considering the same example of Mayan Calendar,  it is possible to compute the single-order-up granularity tun-of-katun  given the multiple-order-up granularity uinal-baktun using equation \ref{eq5}

\begin{equation} \label{eq6}
\begin{split}
C(tun, katun) =  \lfloor C(uinal, baktun)(z)/18\rfloor \mod 20 \\
\end{split}
\end{equation}
